DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life

### 2021-11-02 manuscript/manuscript.Rmd
intro

keywords          : "Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations"
wordcount         : "2650"

@@

>>>>

keywords          : "Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations"
wordcount         : "2949"

@@

Time of evolutionary origin represents a fundamental piece of information for
understanding biological processes in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000], from historical biogeography to species diversification studies [@posadas2006historical; @Morlon2014].
The number of studies publishing phylogenies with branch lengths proportional to absolute time (hereafter chronograms) have constantly increased in number for the last two decades [@Kumar2017].
Still, generating a chronogram is not an easy task unless you have specialized training, and a wealth of time and resources: it requires inferring a phylogenetic tree using genetic markers, obtaining independent time data from the fossil record, and understanding the placement of fossils on the tree as well as their limits for analysis. That is why there has been an urge for promoting and facilitating the reuse of the vast amount of phylogenetic and evolutionary time data that has been made available in publications, for the advantage of research relying on this information [@webb2005phylomatic; @Stoltzfus2013].

A tool for efficient reuse of expert, published data on time of evolutionary origin should have an open and fully public chronogram database storing data in a format suitable for scientific reuse, an automatised way of accessing the information, and straightforward means of comparing and summarizing chronogram information as needed by the user.
A prototype service aiming to meet this criteria was developed over a series of hackathons at the National Evolutionary Synthesis Center [@Stoltzfus2013].
Here we present the full implementation of the `datelife` service, constituted by an R package and a web site available at www.datelife.org/query/.
The current implementation of `datelife` performs the tasks described above. It features an algorithm for automatic curation and maintenance of an open database of chronograms pulled from the OpenTree public repository, methods to summarize and compare source chronograms, and new functions to visualize and graphically compare source and summary chronograms.

### 2021-11-02 manuscript/manuscript.tex
intro

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life},
  pdfauthor={Luna L. Sánchez Reyes1,2, Emily Jane McTavish1, \& Brian O'Meara2},
  pdflang={en-EN},
  pdfkeywords={Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{DateLife: revealing the dated Tree of Life}
\keywords{Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations\newline\indent Word count: 2650}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\abstract{
Time of evolutionary origin is fundamental for research in the natural sciences, as well as for education, science communication and policy.
Despite an increased availability of fossil and molecular data, and time-efficient analytical techniques, achieving a high-quality reconstruction of time of evolutionary origin as a phylogenetic tree with branch lengths proportional to absolute time (chronogram), is still a difficult and time-consuming task for a majority of interested parties.
Yet, the amount of published chronograms has increased significantly in the past two decades, and a non-negligeable proportion of these data have been steadily accumulating in public, open databases such as TreeBASE and Open Tree
of Life, exposing a wealth of expertly-curated and peer-reviewed data on time of evolutionary origin in a programatic and reusable way, for a large quantity and diversity of organisms.
This trend results from intensive and localized efforts for improving data sharing practices, as well as incentivizing open science in biology. Despite these trends, accessibility for time data is not that good.
R has become a widely used element of the biological data analyst toolkit. Hence,to improve accessibility of time data we developed an R package that access these data and eases interaction, reanalysis and reuse of it, incorporation of these data into the evolutionary workflow.

Here we present \texttt{datelife}, a service implemented as an R package and a web site
(www.datelife.org) for increased accessibility, efficient reuse, summary and reanalysis of expert, peer-reviewed, public data on time of evolutionary origin.

Main results:
1. blah
2. blah
3. blah

Results in a general context:
All source and summary chronograms can be saved in formats that permit easy reuse and reanalysis. Summary and newly generated trees are potentially useful to evaluate evolutionary hypothesis in different areas of research in biology.
How well this trees work for this purpose still needs to be tested.

\texttt{datelife} will be useful to increase awereness on the existing variation in expert time of divergence data, and might foster exploration of the effect of alternative divergence time hypothesis on the results of analyses, nurturing a culture of more cautious interpretation of evolutionary results.
}

@@

@@

The main goal of \texttt{datelife} is to generate a chronogram for any given combination of taxon names, based on expert scientific information.

@@

Up to the time this manuscript was written, \texttt{datelife}'s chronogram database had 231 chronograms, pulled entirely from OpenTree's tree repository, the only public tree repository from where \texttt{datelife} can currently get chronograms to construct its database. This represents 5.79\% of the largest existing chronogram database, TimeTree, which has a collection of 3,998 chronograms as of November 01 2021. Unfortunately, TimeTree's database is not open for scientific reuse nor automatised data mining (Kumar, Stecher, Suleski, \& Hedges, 2017).
In 2015, a synthetic chronogram was constructed from 2,274 chronograms available at the time on the TimeTree database (Hedges, Marin, Suleski, Paymer, \& Kumar, 2015). This is the only synthetic TimeTree chronogram that has been made publicly available and deposited on the OpenTree repository, and is part of datelife's database now. Hence, the amount of lineages represented in datelife's database is at least as substantial as TimeTree's, ensuring that some information will be available for any given taxon or lineage. Regrettably, this does not ensure that the full state of knowledge of time of divergence of the taxon/lineage will be available.
Incorporation of more published chronograms into \texttt{datelife}'s database is crucial to improve its services. One option to increase our database is the Dryad data repository. Methods to automatically mine chronograms from Dryad could be designed and implemented. However, Dryad's metadata system has no information to automatically detect branch length units, and those would still need to be determined on a second step, by a curator.
Consequently, we would like to emphasize on the importance of sharing chronogram data for the benefit of the scientific community as a whole, into repositories that require expert input and manual curation, such as OpenTree's tree repository (McTavish et al., 2015).

@@

@@

@@

@@

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life},
  pdfauthor={Luna L. Sánchez Reyes1,2, Emily Jane McTavish1, \& Brian O'Meara2},
  pdflang={en-EN},
  pdfkeywords={Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{DateLife: revealing the dated Tree of Life}
\keywords{Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations\newline\indent Word count: 2949}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\abstract{
Time of evolutionary origin is fundamental for research in the natural sciences, as well as for education, science communication and policy.
Despite an increased availability of fossil and molecular data, and time-efficient analytical techniques, achieving a high-quality reconstruction of time of evolutionary origin as a phylogenetic tree with branch lengths proportional to absolute time (chronogram), is still a difficult and time-consuming task for a majority of interested parties.
Yet, the amount of published chronograms has increased significantly in the past two decades, and a non-negligeable proportion of these data have been steadily accumulating in public, open databases such as TreeBASE and Open Tree
of Life, exposing a wealth of expertly-curated and peer-reviewed data on time of evolutionary origin in a programatic and reusable way, for a large quantity and diversity of organisms.
This trend results from intensive and localized efforts for improving data sharing practices, as well as incentivizing open science in biology. Despite these trends, accessibility to state-of-the-art knowledge on time of evolutionary origin is still reduced.

Here we present \texttt{datelife}, a service implemented as an R package and an Rshiny website application
available at www.datelife.org/query/, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time of evolutionary origin.

The main workflow of \texttt{datelife} is to construct a chronogram for any given combination of taxon names, by searching a local chronogram database constructed and curated from the Open Tree of Life (OpenTree), which incorporates phylogenetic data from the TreeBASE database as well.
We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms. Additionally, time data extracted from source chronograms can be usedas secondary calibration points to add branch lengths proportional to absolute time to a tree topology using alternative dating methods.

Summary and newly generated trees are potentially useful to evaluate evolutionary hypothesis in different areas of research in biology. How well this chronograms work for this purpose still needs to be tested.

\texttt{datelife} will be useful to increase awereness on the existing variation in expert time of divergence data, and might foster exploration of the effect of alternative divergence time hypothesis on the results of analyses, providing a framework for a more informed interpretation of evolutionary results.
}

@@

Time of evolutionary origin represents a fundamental piece of information for
understanding biological processes in many areas of research, from developmental to conservation biology (Felsenstein, 1985; Webb, 2000), from historical biogeography to species diversification studies (Morlon, 2014; Posadas, Crisci, \& Katinas, 2006).
The number of studies publishing phylogenies with branch lengths proportional to absolute time (hereafter chronograms) have constantly increased in number for the last two decades (Kumar, Stecher, Suleski, \& Hedges, 2017).
Still, generating a chronogram is not an easy task unless you have specialized training, and a wealth of time and resources: it requires inferring a phylogenetic tree using genetic markers, obtaining independent time data from the fossil record, and understanding the placement of fossils on the tree as well as their limits for analysis. That is why there has been an urge for promoting and facilitating the reuse of the vast amount of phylogenetic and evolutionary time data that has been made available in publications, for the advantage of research relying on this information (Stoltzfus et al., 2013; Webb \& Donoghue, 2005).

A tool for efficient reuse of expert, published data on time of evolutionary origin should have an open and fully public chronogram database storing data in a format suitable for scientific reuse, an automatised way of accessing the information, and straightforward means of comparing and summarizing chronogram information as needed by the user.
A prototype service aiming to meet this criteria was developed over a series of hackathons at the National Evolutionary Synthesis Center (Stoltzfus et al., 2013).
Here we present the full implementation of the \texttt{datelife} service, constituted by an R package and a web site available at www.datelife.org/query/.
The current implementation of \texttt{datelife} performs the tasks described above. It features an algorithm for automatic curation and maintenance of an open database of chronograms pulled from the OpenTree public repository, methods to summarize and compare source chronograms, and new functions to visualize and graphically compare source and summary chronograms.

@@

@@

Up to the time this manuscript was written, \texttt{datelife}'s chronogram database had 231 chronograms, pulled entirely from OpenTree's tree repository, the only public tree repository from where \texttt{datelife} can currently get chronograms to construct its database. This represents 5.79\% of the largest existing chronogram database, TimeTree, which has a collection of 3,998 chronograms as of November 02 2021. Unfortunately, TimeTree's database is not open for scientific reuse nor automatised data mining (Kumar et al., 2017).
In 2015, a synthetic chronogram was constructed from 2,274 chronograms available at the time on the TimeTree database (Hedges, Marin, Suleski, Paymer, \& Kumar, 2015). This is the only synthetic TimeTree chronogram that has been made publicly available and deposited on the OpenTree repository, and is part of datelife's database now. Hence, the amount of lineages represented in datelife's database is at least as substantial as TimeTree's, ensuring that some information will be available for any given taxon or lineage. Regrettably, this does not ensure that the full state of knowledge of time of divergence of the taxon/lineage will be available.
Incorporation of more published chronograms into \texttt{datelife}'s database is crucial to improve its services. One option to increase our database is the Dryad data repository. Methods to automatically mine chronograms from Dryad could be designed and implemented. However, Dryad's metadata system has no information to automatically detect branch length units, and those would still need to be determined on a second step, by a curator.
Consequently, we would like to emphasize on the importance of sharing chronogram data for the benefit of the scientific community as a whole, into repositories that require expert input and manual curation, such as OpenTree's tree repository (McTavish et al., 2015).

@@

\leavevmode\hypertarget{ref-Felsenstein1985a}{}%
Felsenstein, J. (1985). Phylogenies and the Comparative Method. \emph{The American Naturalist}, \emph{125}(1), 1--15. Retrieved from \url{http://www.jstor.org/stable/2461605}

@@

\leavevmode\hypertarget{ref-Morlon2014}{}%
Morlon, H. (2014). Phylogenetic approaches for studying diversification. \emph{Ecology Letters}, \emph{17}(4), 508--525. \url{https://doi.org/10.1111/ele.12251}

@@

\leavevmode\hypertarget{ref-posadas2006historical}{}%
Posadas, P., Crisci, J. V., \& Katinas, L. (2006). Historical biogeography: A review of its basic concepts and critical issues. \emph{Journal of Arid Environments}, \emph{66}(3), 389--403.

@@

\leavevmode\hypertarget{ref-Stoltzfus2013}{}%
Stoltzfus, A., Lapp, H., Matasci, N., Deus, H., Sidlauskas, B., Zmasek, C. M., \ldots{} Jordan, G. (2013). Phylotastic! Making tree-of-life knowledge accessible, reusable and convenient. \emph{BMC Bioinformatics}, \emph{14}. \url{https://doi.org/10.1186/1471-2105-14-158}

@@

\leavevmode\hypertarget{ref-Webb2000}{}%
Webb, C. O. (2000). Exploring the Phylogenetic Structure of Ecological Communities : An Example for Rain Forest Trees. \emph{The American Naturalist}, \emph{156}(2), 145--155.

@@

\leavevmode\hypertarget{ref-webb2005phylomatic}{}%
Webb, C. O., \& Donoghue, M. J. (2005). Phylomatic: Tree assembly for applied phylogenetics. \emph{Molecular Ecology Notes}, \emph{5}(1), 181--183.

### 2021-11-02 manuscript/paper_3-main.Rmd
intro

Clade ages represent a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000], from historical biogeography to species diversification studies [@posadas2006historical; @Morlon2014]. The primary information needed for these time estimates comes from the fossil record. Coupled with phylogenies with branch lengths based on molecular and/or morphological data, the time of divergence of extant and extinct lineages can be reconstructed with molecular dating methods.
The number of studies publishing phylogenies with branch lengths proportional to geological time (hereafter chronograms) have constantly increased in number for the last two decades [@Kumar2017].
Still, generating a chronogram is not an easy task unless you have specialized training: it requires inferring a tree, understanding what fossil data are available and their limits, and where fossils go on the tree. That is why there has been an urge for promoting and facilitating reuse of the vast amount of phylogenetic and time of lineage divergence data that has been generated and made available in publications, for the advantage of research relying on this information [@webb2005phylomatic; @Stoltzfus2013].

@@

To our knowledge, there is no system allowing the comparison of available source information (apw does this a little bit, but it is a verbal comparison more than a graphical or visual comparison). Other types of technical comparisons: which trees are better (which trees represent reality better), but this is of course a very difficult issue.
 -->
A tool for efficient reuse of expert, published data on time of lineage divergence should have an open and fully public chronogram database storing data in a format suitable for scientific reuse, an automatised way of accessing the information, and straightforward means of comparing and summarizing chronogram information as needed by the user.
A prototype service aiming to meet this characteristics was developed over a series of hackathons at the National Evolutionary Synthesis Center [@Stoltzfus2013].
Here we present the formal description and implementation of the `datelife` service, constituted by an R package and a web site (<http://www.datelife.org/>).
<!-- There is still much room for improvement, and flaws and limitations are addressed below.  -->
The current implementation of `datelife` performs the basic tasks described above. It features a system for maintenance of an open database of chronograms pulled from public repositories, methods to summarize and compare source chronograms, and new functions to visualize and graphically compare source and summary chronograms.
<!-- , for the benefit of the broader community. -->
<!-- as well as tools for comparison of source and summary trees.  NOT YET u_u-->
<!-- R packages for benchmarking of functionalities and demonstrating services were also developed and made available.>

>>>>

 The primary information needed for these time estimates comes from the fossil record. Coupled with phylogenies with branch lengths based on molecular and/or morphological data, the time of divergence of extant and extinct lineages can be reconstructed with molecular dating methods.

@@

To our knowledge, there is no system allowing the comparison of available source information (apw does this a little bit, but it is a verbal comparison more than a graphical or visual comparison). Other types of technical comparisons: which trees are better (which trees represent reality better), but this is of course a very difficult issue.
 -->

<!-- , for the benefit of the broader community. -->
<!-- as well as tools for comparison of source and summary trees.  NOT YET u_u-->
<!-- R packages for benchmarking of functionalities and demonstrating services were also developed and made available.>
